---
title: "Global nematode perspectives"
author: "J. van den Hoogen & S. Geisen et al. 2019"
output:
  html_document:
    # code_folding: hide
    toc: yes
    toc_collapse: no
    toc_float: yes
    theme: spacelab
    fig_retina: 2
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, results='hide', message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(reshape2)
  library(corrplot)
  library(tidyverse)
  library(cowplot)
  library(rworldmap)
  library(NbClust)
  library(factoextra)
  library(cluster)
  library(vegan)
  library(RColorBrewer)
  library(lemon)
  library(parallel)
  library(plotrix)
  library(caret)
  library(scales)
})
```

# About the code
This Rmarkdown file was used to create the main and supplemental figures for the associated paper (Van den Hoogen & Geisen et al., 2019), for which the raw abundance data was used. Scripts used for the geospatial analysis can be found under the respective directory in the main repository. Scripts are fully annotated for and organized by order of appearance in the paper. 

For questions, please send an email to: johan.vandenhoogen@usys.ethz.ch

# Initial data formating
```{r Data import & general calculations}
# Set working directory
setwd("~/Work/ETH/Projects/Nematodes/Data")

# Load data
Data_raw <- read.csv("20190131_NematodePoints_SampledPixelValues_wBiome.csv")

# Add ID column
Data_raw <- tibble::rowid_to_column(Data_raw, "ID") 

# Antartic sites
Antarctica_points <- read.csv("20193119_NematodePoints_Antarctica_SampledPixelValues.csv") 
Antarctica <- Antarctica_points %>% 
  select(-Unidentified, -Pixel_Lat, -Pixel_Long)

# Calculate proportional abundances
Data_raw$Bacterivores.pct <- Data_raw$Bacterivores/rowSums(Data_raw[,c('Total_Number','Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')])
Data_raw$Fungivores.pct <- Data_raw$Fungivores/rowSums(Data_raw[,c('Total_Number','Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')])
Data_raw$Herbivores.pct <- Data_raw$Herbivores/rowSums(Data_raw[,c('Total_Number','Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')])
Data_raw$Omnivores.pct <- Data_raw$Omnivores/rowSums(Data_raw[,c('Total_Number','Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')])
Data_raw$Predators.pct <- Data_raw$Predators/rowSums(Data_raw[,c('Total_Number','Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')])

#select data
Data_tot <- Data_raw[,c('ID','Total_Number','Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')] %>% 
  na.omit() %>% 
  filter(Total_Number != 0)
Data_group <- Data_tot[,c('Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')] %>% 
  na.omit()

#pull used IDs
Used_IDs <- Data_tot[,1,drop=FALSE]
#remove ID column 
Data_tot <- subset(Data_tot, select = -ID)
#convert to data frame
Data_tot <- stack(as.data.frame(Data_tot)) %>% 
    dplyr::rename(Group = ind) %>% 
    dplyr::rename(Count = values) %>% 
    mutate(Count = as.numeric(Count))

#Calculate frequencies per row
GroupComposition <- Data_raw[,c('ID','Bacterivores.pct','Fungivores.pct',"Herbivores.pct",'Omnivores.pct','Predators.pct')]
#remove rows not including all trophic groups
GroupComposition <- merge(GroupComposition, Used_IDs, "ID")
#remove ID column
GroupComposition <- subset(GroupComposition, select = -ID)
names(GroupComposition) <- c('Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')
```

# Figure 1a - Sampling locations
```{r Observations per biome}
# Collate points to new df
total_points <- plyr::rbind.fill(Data_raw, Antarctica_points)

# Define world outline map
world <- map_data("world")

# Define color palette
pal <- brewer.pal(8, "YlOrRd")

# Plot sampling points on map
nematode_pointmap <- ggplot() + geom_polygon(data = world, 
                                             aes(x = long, y = lat, group = group),
                                             fill = "#bababa",
                                             color = NA,
                                             size = 0.1) + 
  coord_fixed(1.1) +
  geom_point(data = total_points,
             aes(x = Pixel_Long, y = Pixel_Lat, fill = Total_Number),
             color = "black",
             pch = 21
  ) +
  scale_fill_gradientn(colors = pal,
                       limits = c(0, 3000),
                       oob = scales::squish,
                       name = "Nematodes per 100g dry soil") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.box="horizontal",
        panel.grid = element_blank(),
        axis.title=element_blank(),
        axis.text=element_blank()) +
  guides(fill = guide_colorbar(title.position = "top"))

nematode_pointmap

# Save to file
ggsave("nematode_pointmap.pdf", plot = nematode_pointmap)
```

# Figure 1b - Observed abundances total and per trophic group, per biome
## Summary per biome
```{r Observations per trophic group and biome}
# create color scheme
col1 <- colorRampPalette(c("#B26A4A","#B2A877","#62B25E","#40B29A","#ef7e6b"))

#plot observations total number
plot_tot <- ggplot(subset(Data_tot, Group == "Total_Number"), aes(
 x = Group, 
 y = Count, 
 # fill = Group,
 color = Group ) ) +
 geom_jitter(size = 0.75, alpha = 0.05, width = 0.35, height = 0) +
 geom_boxplot(outlier.shape = NA, color = "black", alpha = 0.3) + 
 theme_bw() + 
 coord_cartesian(ylim = c(0,8000)) +
 theme(panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    axis.line = element_line(color = 'black'), 
    panel.border = element_rect(color = "white"),
    aspect.ratio = (2.6/1),
    legend.position = "none"
    ) +
 scale_color_manual(values = "#8399B2") + 
  scale_y_continuous(labels = comma) +
 scale_x_discrete(labels = "Total number") +
 ylab("Nematodes per 100 g dry soil")  

#plot observations per trophic group
plot_sub <- ggplot(subset(Data_tot, Group != "Total_Number"), aes(
 x = Group, 
 y = Count, 
 # fill = Group,
 color = Group ) ) +
 geom_jitter(size = 0.75, alpha = 0.05, width = 0.35, height = 0) +
 geom_boxplot(outlier.shape = NA, color = "black", alpha = 0.3) + 
 theme_bw() + 
  coord_fixed(ratio = 1) +
 coord_cartesian(ylim = c(0,3000)) +
 theme(panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    axis.line = element_line(color = 'black'), 
    panel.border = element_rect(color = "white"),
    legend.position = "none") +
 scale_fill_manual(values = col1(5)) +
 scale_color_manual(values = col1(5)) + 
 xlab("") +
 ylab("") +
  scale_y_continuous(labels = comma)

# Data formating
setwd("~/Work/ETH/Projects/Nematodes/Data/")

Data_biome <- Data_raw %>% 
  select('Total_Number','Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators','WWF_Biome') %>% 
  plyr::rbind.fill(Antarctica) %>%
  filter(WWF_Biome != 2) %>% 
  filter(WWF_Biome != 3) %>% 
  filter(WWF_Biome != 9) %>% 
  filter(WWF_Biome != 98) %>% 
  na.omit() %>% 
  melt(id.vars = "WWF_Biome") %>%
  mutate(WWF_Biome = replace(WWF_Biome, WWF_Biome == 1, "Tropical Moist Forests")) %>%
  mutate(WWF_Biome = replace(WWF_Biome, WWF_Biome == 4, "Temperate Broadleaf Forests")) %>%
  mutate(WWF_Biome = replace(WWF_Biome, WWF_Biome == 5, "Temperate Conifer Forests")) %>%
  mutate(WWF_Biome = replace(WWF_Biome, WWF_Biome == 6, "Boreal Forests")) %>%
  mutate(WWF_Biome = replace(WWF_Biome, WWF_Biome == 7, "Tropical Grasslands")) %>%
  mutate(WWF_Biome = replace(WWF_Biome, WWF_Biome == 8, "Temperate Grasslands")) %>%
  mutate(WWF_Biome = replace(WWF_Biome, WWF_Biome == 10, "Montane Grasslands")) %>%
  mutate(WWF_Biome = replace(WWF_Biome, WWF_Biome == 11, "Tundra")) %>%
  mutate(WWF_Biome = replace(WWF_Biome, WWF_Biome == 12, "Mediterranean Forests")) %>%
  mutate(WWF_Biome = replace(WWF_Biome, WWF_Biome == 13, "Deserts")) %>% 
  mutate(WWF_Biome = replace(WWF_Biome, WWF_Biome == 25, "Antarctica")) %>% 
  setNames(., c("Biome","Group","Count")) 

Data_biome_sum <- Data_biome %>%
  filter(Group == "Total_Number") %>% 
  group_by(Biome) %>% 
  dplyr::summarize(Mean = mean(Count, na.rm=TRUE), 
                   Median = median(Count, na.rm=TRUE),
                   n = n()) %>% 
  arrange(desc(Median))  

Data_biome_sum
write.csv(Data_biome_sum, "data_biome_sum.csv")

Data_biome_tot <- Data_biome %>% filter(Group == "Total_Number")

# ordering of boxplots per biome. Ordered by median of total number in each biome
levels = levels(reorder(factor(Data_biome_tot$Biome), Data_biome_tot$Count, median))
```

## Summary per trophic group
Note: to create the final figure, we made some final modifications to the produced plot was modified in Adobe Illustrator CC 2018.
```{r Trophic group and per biome, message = FALSE}
# Creating subplots
Total <- ggplot(data = Data_biome_tot, aes(
  x = factor(Biome, levels = levels),
  y = Count,
  color = factor(Biome),
  alpha = 0.2) ) + 
  geom_jitter(alpha = 0.1, size = 0.5, color = "#8399B2") +
  geom_boxplot(outlier.shape = NA, color = "black", fill = NA, lwd = 0.35) +
  theme_minimal() +
  scale_alpha(guide = 'none') +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.title.y=element_blank(),
    axis.title.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.x=element_line(),
    plot.title = element_text(hjust = 0.5, size = 10),
    legend.position = "none",
    axis.text.x = element_text(angle = 60, hjust = 1)) +
  xlab("Nematodes per 100 g dry soil)") +
  coord_flip(ylim = c(0,10000)) + 
  scale_y_continuous(breaks = c(0, 5000, 9999), labels = comma) # Max value of 9,999 as workaround, to prevent scaling issues for final figure

Bacterivores <- ggplot(data = subset(Data_biome, Group == "Bacterivores"), aes(
  x = factor(Biome, levels = levels),
  y = Count,
  color = factor(Biome),
  alpha = 0.2) ) + 
  geom_jitter(alpha = 0.1, size = 0.5, color = "#B26A4A") +
  geom_boxplot(outlier.shape = NA, color = "black", fill = NA, lwd = 0.35) +
  theme_minimal() +
  scale_alpha(guide = 'none') +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.title.y=element_blank(),
    axis.title.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.x=element_line(),
    plot.title = element_text(hjust = 0.5, size = 10),
    legend.position = "none",
    axis.text.x = element_text(angle = 60, hjust = 1)) +
  xlab("Nematodes per 100 g dry soil)") +
  coord_flip(ylim = c(0,4000)) +
  scale_y_continuous(breaks = c(0,2000, 4000), labels = comma)

Fungivores <- ggplot(data = subset(Data_biome, Group == "Fungivores"), aes(
  x = factor(Biome, levels = levels),
  y = Count,
  alpha = 0.2) ) + 
  geom_jitter(alpha = 0.1, size = 0.5, color = "#B2A877") +
  geom_boxplot(outlier.shape = NA, color = "black", fill = NA, lwd = 0.35) +
  theme_minimal() +
  scale_alpha(guide = 'none') +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.title.y=element_blank(),
    axis.title.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.x=element_line(),
    plot.title = element_text(hjust = 0.5, size = 10),
    legend.position = "none",
    axis.text.x = element_text(angle = 60, hjust = 1)) +
  xlab("") +
  coord_flip(ylim = c(0,2000)) +
  scale_y_continuous(breaks = c(0, 1000, 2000), labels = comma)

Herbivores <- ggplot(data = subset(Data_biome, Group == "Herbivores"), aes(
  x = factor(Biome, levels = levels),
  y = Count,
  alpha = 0.2) ) + 
  geom_jitter(alpha = 0.1, size = 0.5, color = "#62B25E") +
  geom_boxplot(outlier.shape = NA, color = "black", fill = NA, lwd = 0.35) +
  theme_minimal() +
  scale_alpha(guide = 'none') +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.title.y=element_blank(),
    axis.title.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.x=element_line(),
    plot.title = element_text(hjust = 0.5, size = 10),
    legend.position = "none",
    axis.text.x = element_text(angle = 60, hjust = 1)) +
  xlab("") +
  coord_flip(ylim = c(0,4000)) +
  scale_y_continuous(breaks = c(0,2000, 4000), labels = comma)

Omnivores <- ggplot(data = subset(Data_biome, Group == "Omnivores"), aes(
  x = factor(Biome, levels = levels),
  y = Count,
  alpha = 0.2) ) + 
  geom_jitter(alpha = 0.1, size = 0.5, color ="#40B29A") +
  geom_boxplot(outlier.shape = NA, color = "black", fill = NA, lwd = 0.35) +
  theme_minimal() +
  scale_alpha(guide = 'none') +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.title.y=element_blank(),
    axis.title.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.x=element_line(),
    plot.title = element_text(hjust = 0.5, size = 10),
    legend.position = "none",
    axis.text.x = element_text(angle = 60, hjust = 1)) +
  xlab("") +
  coord_flip(ylim = c(0,800)) +
  scale_y_continuous(breaks = c(0,400, 800), labels = comma)

Predators <- ggplot(data = subset(Data_biome, Group == "Predators"), aes(
  x = factor(Biome, levels = levels),
  y = Count,
  alpha = 0.2) ) + 
  geom_jitter(alpha = 0.1, size = 0.5, color = "#ef7e6b") +
  geom_boxplot(outlier.shape = NA, color = "black", fill = NA, lwd = 0.35) +
  theme_minimal() +
  scale_alpha(guide = 'none') +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.title.y=element_blank(),
    axis.title.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.x=element_line(),
    plot.title = element_text(hjust = 0.5, size = 10),
    legend.position = "none",
    axis.text.x = element_text(angle = 60, hjust = 1)) +
  xlab("") +
  coord_flip(ylim = c(0,300)) +
  scale_y_continuous(breaks = c(0,150, 300), labels = comma)

# Create empty plot witgh only axis labels 
empty <- ggplot(data = subset(Data_biome, Group == "Total_Number"), aes(
  x = factor(Biome, levels = levels),
  Count, 
  alpha = 0.2 )) +
  theme_minimal() +
  scale_alpha(guide = 'none') +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    legend.position = "none") +
  coord_flip() 

# Combine plots to semi-final plot. Adobe Illustrator CC 2018 was used for minor tweaks. 
grid1 <- plot_grid(Bacterivores, Fungivores, Herbivores, Omnivores, Predators, align = 'h', nrow = 1)
plot_obs <- plot_grid(plot_grid(plot_tot, empty, align = "h", ncol = 1), plot_grid(plot_sub, grid1, align = "v", ncol = 1), nrow = 1, rel_widths = c(1,1.3))
plot_obs

gridx <- plot_grid(Total, Fungivores, Herbivores, Omnivores, Predators, align = 'h', nrow = 1)
plot_obs_x <- plot_grid(plot_grid(plot_tot, empty, align = "h", ncol = 1), plot_grid(plot_sub, gridx, align = "v", ncol = 1), nrow = 1, rel_widths = c(1,1.3))
plot_obs_x

# Save to file
setwd("~/Work/ETH/Projects/Nematodes/Figures/")
ggsave("Abundances_biome.pdf", plot = plot_obs, device = "pdf", dpi = "retina", scale = 1)

setwd("~/Work/ETH/Projects/Nematodes/Figures/")
ggsave("Abundances_biome_x.pdf", plot = plot_obs_x, device = "pdf", dpi = "retina", scale = 1)
```

# Figure 2 - Validation
## Figure 2a - standard error observations
```{r Fig 2a - stdev, message=FALSE, warning=FALSE}
# Format data
Data_tot <- Data_raw[,c('Total_Number','Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')] %>% 
  na.omit() %>% 
  filter(Total_Number != 0) %>% 
  dplyr::rename("Total number" = Total_Number)

Data_tot <- stack(as.data.frame(Data_tot)) %>% 
    dplyr::rename(Group = ind) %>% 
    dplyr::rename(Count = values) %>% 
    group_by(Group) 

#number of random samples to be selected
nsamples = c(10,25,50,75,100,150,250,500)
#create empty lists for storing samples
random_samples <- list()
list <- list()
se_list <- list()
  
for (s in 1:100){
  
#select random samples
 for (i in nsamples){
    #set seed
    set.seed(s)
   
    #select random samples 
    random_samples[[i]] <- sample_n(Data_tot, i, replace = T)
    
    #calculate std error, store in list
    list[[i]] <- dplyr::summarise(group_by(as.data.frame(random_samples[i]), Group), se = std.error(Count))
    }
  
 #store std errors per nsamples in df
 se_df <- Reduce(function(x, y) merge(x, y, by = "Group", all = TRUE) ,list(as.data.frame(list[10]),
                                                                             as.data.frame(list[25]),
                                                                             as.data.frame(list[50]),
                                                                             as.data.frame(list[75]),
                                                                             as.data.frame(list[100]),
                                                                             as.data.frame(list[150]),
                                                                             as.data.frame(list[250]),
                                                                             as.data.frame(list[500])))

 #rename columns, melt data
 names(se_df) <- c("Group", "5", "25", "50", "75", "100", "150", "250","500")
 se_df <- melt(se_df, id = "Group")
 names(se_df) <- c("Group", "nsamples", "se")
 #store in list
 se_list[[s]] <- se_df
}

# Function to merge data frames from list
merge.all <- function(x, y) {merge(x, y, all = TRUE, by = c("Group","nsamples"))}
output <- Reduce(merge.all, se_list)
output$mean.se <- rowMeans(output[,3:102])
output <- output[,c("Group","nsamples","mean.se")]
output$nsamples <- as.numeric(as.character(output$nsamples)) 
  
col_o <- colorRampPalette(c("#8399B2","#B26A4A","#B2A877","#62B25E","#40B29A","#ef7e6b"))

# Plot
se_plot_o <- ggplot(output, aes(
  x = nsamples,
  y = mean.se,
  group = Group,
  color = factor(Group))) + 
  geom_point() +
  geom_smooth(se = F, method = "loess", formula = y ~ log(x)) +
  scale_color_manual(values = col_o(6), name = "Trophic group") +
  xlab("Sample size") +
  ylab("Standard error \n (nematodes per 100 g dry soil)") +
  scale_y_continuous(labels = comma, breaks = c(0,250,500,750,1000)) 

se_plot_o

# Save to file
setwd("~/Work/ETH/Projects/Nematodes/Figures/")
ggsave("stderr_plot_obs.pdf", plot = se_plot_o, device = "pdf", dpi = "retina", scale = 2)  
```

## Figure 2b - standard error predicted points
```{r Figure 2b - standard error predicted points, message=FALSE, warning=FALSE}
# select data
setwd("~/Work/ETH/Projects/Nematodes/Calculations/")
std_err_pred <- read.csv("20180731_Nematode_BootStrap_StdError_1000Seeds.csv") %>%
  dplyr::rename(nsamples = NumOfSamples) %>% 
  dplyr::rename(mean.se = StdError) %>% 
  mutate(Biome = replace(Biome, Biome == 1, "Tropical Moist Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 2, "Tropical Dry Forest")) %>%
  mutate(Biome = replace(Biome, Biome == 3, "Tropical Coniferous Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 4, "Temperate Broadleaf Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 5, "Temperate Conifer Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 6, "Boreal Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 7, "Tropical Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 8, "Temperate Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 9, "Flooded Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 10, "Montane Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 11, "Tundra")) %>%
  mutate(Biome = replace(Biome, Biome == 12, "Mediterranean Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 13, "Deserts")) %>% 
  mutate(Biome = replace(Biome, Biome == 14, "Mangroves")) 

levels <- c("Tropical Moist Forests",
            "Tropical Dry Forest",
            "Tropical Coniferous Forests",
            "Temperate Broadleaf Forests",
            "Temperate Conifer Forests",
            "Boreal Forests",
            "Tropical Grasslands",
            "Temperate Grasslands",
            "Flooded Grasslands",
            "Montane Grasslands",
            "Tundra",
            "Mediterranean Forests",
            "Deserts",
            "Mangroves")

# Set color scheme
col <- colorRampPalette(brewer.pal(12, "Accent"))

#plot
se_plot_p <- ggplot(std_err_pred, aes(
    x = nsamples,
    y = mean.se,
    group = Biome,
    color = factor(Biome))) + 
    geom_point() +
  geom_smooth(se = F, method = "loess", formula = y ~ log(x)) +
    scale_color_manual(values = col(14),
                       name = "Biome",
                       breaks = levels) +
    xlab("Sample size") +
    ylab("Standard error \n (nematodes per 100 g dry soil)")

  
se_plot_p

# save to file
setwd("~/Work/ETH/Projects/Nematodes/Figures/")
ggsave("stderr_plot_pred.pdf", plot = se_plot_p, device = "pdf", dpi = "retina", scale = 2)  
```

## Final Figure 2
To create the final figures, some minor modifications were made using Adobe Illustrator CC 2018.
```{r}
# Combine standard error plots
se_plot_o <- se_plot_o + theme(legend.position = c(0.5, 0.7), 
                               legend.text = element_text(size = 10),
                               # legend.title = element_text(size = 12),
                               legend.key = element_rect(size = 1),
                               legend.key.size = unit(0.65, 'lines'),
                               legend.title = element_blank())
se_plot_p <- se_plot_p + theme(legend.position = c(0.3, 0.7), 
                               legend.text = element_text(size = 10),
                               # legend.title = element_text(size = 12),
                               # axis.title.y = element_blank(),
                               legend.key = element_rect(size = 1),
                               legend.key.size = unit(0.65, 'lines'),
                               legend.title = element_blank())

se_plot_combined <- plot_grid(se_plot_o, se_plot_p, align = "h", labels = "auto")
se_plot_combined

# Save to file
setwd("~/Work/ETH/Projects/Nematodes/Figures/")
save_plot("stderr_plot_combined.pdf", se_plot_combined, base_aspect_ratio = 2.2)
# ggsave("stderr_plot_combined.pdf", plot = se_plot_combined, device = "pdf")  

```

# Figures 2c-h - Predicted vs Observed
```{r Predicted vs Observed plots}
setwd("/Users/johanvandenhoogen/work/ETH/projects/Nematodes/R_code/20190607_Johan_Predicted_Versus_Observed/")

# Import the raw data
Total_Number <- read.csv('20180723_TotalNumb_PredVsObs.csv') %>% 
  select(-system.index, -.geo) %>% 
  rename(Observed = Total_Numb, Predicted = classification)
Bacterivores <- read.csv('20180723_Bacterivores_PredVsObs.csv') %>% 
  select(-system.index, -.geo) %>% 
  rename(Observed = Bacterivor, Predicted = classification)
Fungivores <- read.csv('20180723_Fungivores_PredVsObs.csv') %>% 
  select(-system.index, -.geo) %>% 
  rename(Observed = Fungivores, Predicted = classification)
Herbivores <- read.csv('20180723_Herbivores_PredVsObs.csv') %>% 
  select(-system.index, -.geo) %>% 
  rename(Observed = Plant_feed, Predicted = classification)
Omnivores <- read.csv('20180723_Omnivores_PredVsObs.csv') %>% 
  select(-system.index, -.geo) %>% 
  rename(Observed = Omnivores, Predicted = classification)
Predators <- read.csv('20180723_Predators_PredVsObs.csv') %>% 
  select(-system.index, -.geo) %>% 
  rename(Observed = Predators, Predicted = classification)

# Define palette
paletteForUse <- c('#d10000', '#ff6622', '#ffda21', '#33dd00', '#1133cc', '#220066', '#330044')
colors <-  colorRampPalette(paletteForUse)(256)

# Total_Number
# Use densCols() output to get density at each point
Total_Number$dens <- col2rgb(densCols(Total_Number$Observed, Total_Number$Predicted))[1,] + 1L

# Calculate predicted - observed R2
R2Total_Number <- round(1 - (sum((Total_Number$Observed-Total_Number$Predicted )^2)/sum((Total_Number$Observed-mean(Total_Number$Observed))^2)),2)

# Map densities to colors
Total_Number$colors = colors[Total_Number$dens]

total_number_plot <- Total_Number %>% 
  ggplot(aes(x = Observed, y = Predicted)) +
  geom_point(color = Total_Number$colors) +
  coord_cartesian(xlim = c(0,15000), ylim = c(0,15000)) +
  geom_smooth(method = 'lm', se = FALSE, linetype = 'dashed', color = 'black') + 
  geom_abline() +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  theme_bw() +
  theme(aspect.ratio = 1,
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Total Number")

# Bacterivores
# Use densCols() output to get density at each point
Bacterivores$dens <- col2rgb(densCols(Bacterivores$Observed, Bacterivores$Predicted))[1,] + 1L

# Calculate predicted - observed R2
R2Bacterivores <- round(1 - (sum((Bacterivores$Observed-Bacterivores$Predicted )^2)/sum((Bacterivores$Observed-mean(Bacterivores$Observed))^2)),2)

# Map densities to colors
Bacterivores$colors = colors[Bacterivores$dens]

bacterivores_plot <- Bacterivores %>% 
  ggplot(aes(x = Observed, y = Predicted)) +
  geom_point(color = Bacterivores$colors) +
  coord_cartesian(xlim = c(0,10000), ylim = c(0,10000)) +
  geom_smooth(method = 'lm', se = FALSE, linetype = 'dashed', color = 'black') + 
   geom_abline() +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  theme_bw() +
  theme(aspect.ratio = 1,
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Bacterivores")

# Fungivores
# Use densCols() output to get density at each point
Fungivores$dens <- col2rgb(densCols(Fungivores$Observed, Fungivores$Predicted))[1,] + 1L

# Calculate predicted - observed R2
R2Fungivores <- round(1 - (sum((Fungivores$Observed-Fungivores$Predicted )^2)/sum((Fungivores$Observed-mean(Fungivores$Observed))^2)),2)

# Map densities to colors
Fungivores$colors = colors[Fungivores$dens]

fungivores_plot <- Fungivores %>% 
  ggplot(aes(x = Observed, y = Predicted)) +
  geom_point(color = Fungivores$colors) +
  coord_cartesian(xlim = c(0,2000), ylim = c(0,2000)) +
  geom_smooth(method = 'lm', se = FALSE, linetype = 'dashed', color = 'black') + 
  geom_abline() +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  theme_bw() +
  theme(aspect.ratio = 1,
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Fungivores")

# Herbivores
# Use densCols() output to get density at each point
Herbivores$dens <- col2rgb(densCols(Herbivores$Observed, Herbivores$Predicted))[1,] + 1L

# Calculate predicted - observed R2
R2Herbivores <- round(1 - (sum((Herbivores$Observed-Herbivores$Predicted )^2)/sum((Herbivores$Observed-mean(Herbivores$Observed))^2)),2)

# Map densities to colors
Herbivores$colors = colors[Herbivores$dens]

herbivores_plot <- Herbivores %>% 
  ggplot(aes(x = Observed, y = Predicted)) +
  geom_point(color = Herbivores$colors) +
  coord_cartesian(xlim = c(0,5000), ylim = c(0,5000)) +
  geom_smooth(method = 'lm', se = FALSE, linetype = 'dashed', color = 'black') + 
  geom_abline() +
  scale_x_continuous(labels = comma, breaks = c(0,2500, 5000)) +
  scale_y_continuous(labels = comma, breaks = c(0,2500, 5000)) +
  theme_bw() +
  theme(aspect.ratio = 1,
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Herbivores")

# Omnivores
# Use densCols() output to get density at each point
Omnivores$dens <- col2rgb(densCols(Omnivores$Observed, Omnivores$Predicted))[1,] + 1L

# Calculate predicted - observed R2
R2Omnivores <- round(1 - (sum((Omnivores$Observed-Omnivores$Predicted )^2)/sum((Omnivores$Observed-mean(Omnivores$Observed))^2)),2)

# Map densities to colors
Omnivores$colors = colors[Omnivores$dens]

omnivores_plot <- Omnivores %>% 
  ggplot(aes(x = Observed, y = Predicted)) +
  geom_point(color = Omnivores$colors) +
  coord_cartesian(xlim = c(0,1500), ylim = c(0,1500)) +
  geom_smooth(method = 'lm', se = FALSE, linetype = 'dashed', color = 'black') + 
  geom_abline() +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  theme_bw() +
  theme(aspect.ratio = 1,
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Omnivores")

# Predators
# Use densCols() output to get density at each point
Predators$dens <- col2rgb(densCols(Predators$Observed, Predators$Predicted))[1,] + 1L

# Calculate predicted - observed R2
R2Predators <- round(1 - (sum((Predators$Observed-Predators$Predicted )^2)/sum((Predators$Observed-mean(Predators$Observed))^2)),2)

# Map densities to colors
Predators$colors = colors[Predators$dens]

predators_plot <- Predators %>% 
  ggplot(aes(x = Observed, y = Predicted)) +
  geom_point(color = Predators$colors) +
  coord_cartesian(xlim = c(0,1500), ylim = c(0,1500)) +
  geom_smooth(method = 'lm', se = FALSE, linetype = 'dashed', color = 'black') + 
  geom_abline() +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  theme_bw() +
  theme(aspect.ratio = 1,
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "Predators")

# Save to file
setwd("~/Work/ETH/Projects/Nematodes/Figures/")

ggsave("pred_obs_total_number.pdf", total_number_plot, width = 8, height = 8, device = "pdf")
ggsave("pred_obs_bacterivores.pdf", bacterivores_plot, width = 4, height = 4, device = "pdf")
ggsave("pred_obs_fungivores.pdf", fungivores_plot, width = 4, height = 4, device = "pdf")
ggsave("pred_obs_herbivores.pdf", herbivores_plot, width = 4, height = 4, device = "pdf")
ggsave("pred_obs_omnivores.pdf", omnivores_plot, width = 4, height = 4, device = "pdf")
ggsave("pred_obs_predators.pdf", predators_plot, width = 4, height = 4, device = "pdf")

plot_grid(total_number_plot,bacterivores_plot,fungivores_plot,herbivores_plot,omnivores_plot,predators_plot)

p1 <- plot_grid(bacterivores_plot,fungivores_plot, ncol = 1)
p2 <- plot_grid(total_number_plot, p1, rel_widths = c(2,1))
p3 <- plot_grid(herbivores_plot,omnivores_plot,predators_plot, nrow = 1, align = "h")

plot_grid(p2, p3, ncol = 1, rel_heights = c(3,1))

# Legend with color scale
pdf(file='Nema_PredVsObs_Legend_Spectrum.pdf')
legend_image = as.raster(matrix(colors, ncol=1))
plot(c(-3,4),c(0,1),type = 'n', axes = F,xlab = '', ylab = '')
# text(x=1.5, y = seq(0,1,l=5), labels = c('0','','','','1'))
rasterImage(legend_image, 0, 0, 1,1)
dev.off()
```

# Extended Data Figure 2 - Linear regression models
```{r Regression coefficients}
# # Select top 10 variables from RF model
lm_vars <- merge(Used_IDs, Data_raw, by = "ID" ) %>% select(Total_Number,
                                                            Sand_Content_15cm,
                                                            OrgCStockTHa_5to15cm,
                                                            CatIonExcCap_15cm,
                                                            Nadir_Reflectance_Band1,
                                                            Nadir_Reflectance_Band7,
                                                            Annual_Precipitation,
                                                            pHinHOX_15cm,
                                                            Aridity_Index,
                                                            Temperature_Seasonality,
                                                            Precipitation_Seasonality)

# Scale data
lm_vars_scaled <- merge(as.data.frame(lm_vars$Total_Number) %>% tibble::rowid_to_column("ID"),
                        as.data.frame(scale(lm_vars[, -c(1)], center = TRUE, scale = TRUE)) %>% tibble::rowid_to_column("ID"),
                        by = "ID") %>%
  select(-ID) %>%
  dplyr::rename(Total_Number = "lm_vars$Total_Number")

lm_vars_scaled <- melt(lm_vars_scaled, "Total_Number")
# lm_vars <- melt(lm_vars, "Total_Number")

# Source of "stat_smooth_func" and "StatSmoothFunc": https://gist.github.com/kdauria/524eade46135f6348140
# Slightly modified 
{stat_smooth_func <- function(mapping = NULL, data = NULL,
                             geom = "smooth", position = "identity",
                             ...,
                             method = "auto",
                             formula = y ~ x,
                             se = TRUE,
                             n = 80,
                             span = 0.75,
                             fullrange = FALSE,
                             level = 0.95,
                             method.args = list(),
                             na.rm = FALSE,
                             show.legend = NA,
                             inherit.aes = TRUE,
                             xpos = NULL,
                             ypos = NULL) {
  layer(
    data = data,
    mapping = mapping,
    stat = StatSmoothFunc,
    geom = geom,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      method = method,
      formula = formula,
      se = se,
      n = n,
      fullrange = fullrange,
      level = level,
      na.rm = na.rm,
      method.args = method.args,
      span = span,
      xpos = xpos,
      ypos = ypos,
      ...
    )
  )
}

StatSmoothFunc <- ggproto("StatSmooth", Stat,
                          
                          setup_params = function(data, params) {
                            # Figure out what type of smoothing to do: loess for small datasets,
                            # gam with a cubic regression basis for large data
                            # This is based on the size of the _largest_ group.
                            if (identical(params$method, "auto")) {
                              max_group <- max(table(data$group))
                              
                              if (max_group < 1000) {
                                params$method <- "loess"
                              } else {
                                params$method <- "gam"
                                params$formula <- y ~ s(x, bs = "cs")
                              }
                            }
                            if (identical(params$method, "gam")) {
                              params$method <- mgcv::gam
                            }
                            
                            params
                          },
                          
                          compute_group = function(data, scales, method = "auto", formula = y~x,
                                                   se = TRUE, n = 80, span = 0.75, fullrange = FALSE,
                                                   xseq = NULL, level = 0.95, method.args = list(),
                                                   na.rm = FALSE, xpos=NULL, ypos=NULL) {
                            if (length(unique(data$x)) < 2) {
                              # Not enough data to perform fit
                              return(data.frame())
                            }
                            
                            if (is.null(data$weight)) data$weight <- 1
                            
                            if (is.null(xseq)) {
                              if (is.integer(data$x)) {
                                if (fullrange) {
                                  xseq <- scales$x$dimension()
                                } else {
                                  xseq <- sort(unique(data$x))
                                }
                              } else {
                                if (fullrange) {
                                  range <- scales$x$dimension()
                                } else {
                                  range <- range(data$x, na.rm = TRUE)
                                }
                                xseq <- seq(range[1], range[2], length.out = n)
                              }
                            }
                            # Special case span because it's the most commonly used model argument
                            if (identical(method, "loess")) {
                              method.args$span <- span
                            }
                            
                            if (is.character(method)) method <- match.fun(method)
                            
                            base.args <- list(quote(formula), data = quote(data), weights = quote(weight))
                            model <- do.call(method, c(base.args, method.args))
                            
                            m = model
                            eq <- substitute(italic(R)^2~"="~r2, 
                                             list(a = format(coef(m)[1], digits = 3), 
                                                  b = format(coef(m)[2], digits = 3), 
                                                  r2 = format(summary(m)$r.squared, digits = 3)))
                            func_string = as.character(as.expression(eq))
                            
                            if(is.null(xpos)) xpos = min(data$x)*0.9
                            if(is.null(ypos)) ypos = max(data$y)*0.9
                            data.frame(x=xpos, y=ypos, label=func_string)
                            
                          },
                          
                          required_aes = c("x", "y")
)
}

# Plot
lm_plot <- ggplot(data = lm_vars_scaled, aes(
  x = value, 
  y = Total_Number, 
  label = Total_Number)) +
  stat_smooth_func(geom = "text", method = "lm", hjust = 0, parse = TRUE) +
  geom_point(aes(color = variable, alpha = 0.05)) + 
  geom_smooth(method = "lm", se = 0.95) +
  facet_wrap(~variable, scales = "free_x", ncol = 5) +
  theme(legend.position="none", 
        aspect.ratio = 1) +
  xlab("Value") +
  ylab("Total number")

lm_plot

setwd("~/Work/ETH/Projects/Nematodes/Figures/")
ggsave("lm_plot.pdf", plot = lm_plot, device = "pdf", dpi = "retina", scale = 3)
```

# Extended Data Figure 6
## Extended Data Figure 6a - functional group correlation
Calculate correlation for each trophic group.
```{r Figure S6 - trophic group correlation, message=FALSE, warning=FALSE}
# calculate correlation values
NemaCor <- cor(Data_group, use = "pairwise.complete.obs", method = "spearman")

#set colors scale
col <- colorRampPalette(c('#330044', '#220066', '#1133cc', '#33dd00', '#ffda21', '#ff6622', '#d10000'))

# plot 
corrplot(NemaCor, 
         method = "circle",
         type = "upper", 
         tl.col = "black", 
         tl.srt = 45, 
         addgrid.col = NA,
         diag = FALSE,
         col = col(1000))

#save to file
setwd("~/Work/ETH/Projects/Nematodes/Figures/")
dev.print(pdf, "Correlation.pdf")   
```

## Extended Data Figures 6b, c - community composition
Here we explored the soil nematode community composition on the global scale. To do so, we first cluster the communities (i.e., the relative abundance of each of the functional groups in one 1-km2 pixel) into the most representative. Then, using the covariate data, we produce a NMDS plot to explore the importance and directionality of the different environmental factors. As covariate layer information is not available for Antarctic points, those points are removed from the dataset for the NMDS and linear modeling.

```{r NMDS dataset}
# Data_raw <- Data_raw %>% filter(WWF_Biome == 999)

#select data
Data_tot <- Data_raw[,c('ID','Total_Number','Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')] %>% 
  na.omit() %>% 
  filter(Total_Number != 0)
Data_group <- Data_tot[,c('Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')] %>% 
  na.omit()

#pull used IDs
Used_IDs <- Data_tot[,1,drop=FALSE]
#remove ID column 
Data_tot <- subset(Data_tot, select = -ID)
#convert to data frame
Data_tot <- stack(as.data.frame(Data_tot)) %>% 
    dplyr::rename(Group = ind) %>% 
    dplyr::rename(Count = values) %>% 
    mutate(Count = as.numeric(Count))

#Calculate frequencies per row
GroupComposition <- Data_raw[,c('ID','Bacterivores.pct','Fungivores.pct',"Herbivores.pct",'Omnivores.pct','Predators.pct')]
#remove rows not including all trophic groups
GroupComposition <- merge(GroupComposition, Used_IDs, "ID")
#remove ID column
GroupComposition <- subset(GroupComposition, select = -ID)
names(GroupComposition) <- c('Bacterivores','Fungivores',"Herbivores",'Omnivores','Predators')
```

## Non-metric multidimensional scaling
First, we selected the number of clusters to use to cluster the observed soil nematode communities. Based on k-mediods clustering and pairwise distances we chose to select four clusters. 

```{r clustering}
# Partitioning Around Medoids (PAM) - k-mediods clustering
set.seed(123)
pam.res <- pam(GroupComposition, 4)
```

## Community type plots
Then we plotted the community composition to visualize the relative presence of each trophic group.
```{r Plotting the community}
# Pull cluster info per id
cluster_info <- as.data.frame(pam.res$cluster)
#add IDs
cluster_info <- tibble::rowid_to_column(cluster_info, "ID")
names(cluster_info) <- c("ID","Cluster")

GroupCompositionID <- tibble::rowid_to_column(GroupComposition, "ID")
Nema_group_clusters <- merge(GroupCompositionID, cluster_info, by="ID")
Nema_group_clusters <- melt(Nema_group_clusters, id = c("ID","Cluster"))
names(Nema_group_clusters) <- c("ID","Cluster","Group","Freq")

col <- colorRampPalette(c("#e24646", "#bb6cc9", "#11c63f", "#c68311", "#1186c6"))
col1 <- colorRampPalette(c("#B26A4A","#B2A877","#62B25E","#40B29A","#ef7e6b"))

#plot
plots <- ggplot(data = Nema_group_clusters, aes(Freq, fill = Group, alpha = 0.3)) +
  geom_density() +
  scale_alpha(guide = 'none') +
  guides(fill = guide_legend(override.aes= list(alpha = 0.5))) +
  theme_minimal() +
  xlab("Frequency") +
  ylab("") +
  scale_fill_manual(values = col(5),
                      name = "Trophic group",
                      labels = c("Bacterivores","Fungivores","Herbivores", "Omnivores ", "Predators")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y=element_blank(),
        aspect.ratio = 1,
        axis.line = element_line() ) +
  xlim(0,0.5) +
  facet_rep_wrap(~Cluster,repeat.tick.labels = TRUE)
plots

##GroupComposition <- subset(GroupComposition, select = -ID)
#save to file 
setwd("~/Work/ETH/Projects/Nematodes/Figures/")
ggsave("comm_plots.pdf", plot = plots, device = "pdf", height =  5, width = 5, units = "cm", dpi = "retina", scale = 2)    
```

## NMDS
Next, we ran a non-metric multidimensional scaling analysis. 
```{r NMDS}
### The NMDS takes a while to perform with 500 runs. 
### To reduce rendering time of the script the analysis is now commented and a stored .rda file is imported. 

# #Set number of clusters
# clus <- makeCluster(3)
# #random seed
# set.seed(123)
# #NMDS calc
# NemaNMDS <- metaMDS(comm = GroupComposition,
#                     distance = "bray",
#                     k = 3,
#                     pc = TRUE,
#                     autotransform = FALSE,
#                     trymax = 500,
#                     center = TRUE,
#                     parallel = clus
#                     )
# #stop cluster
# stopCluster(clus)
# 
# # save rda file
# setwd("~/Work/ETH/Projects/Nematodes/Data/")
# #Save to file
# save(NemaNMDS, file = "NemaNMDS_31012019.rda")

# # load rda file from disk
setwd("~/Work/ETH/Projects/Nematodes/Data/")
load(file = "NemaNMDS_31012019.rda")

# Plot stressplot
stressplot(NemaNMDS, p.col = "grey", l.col = "black")

# save to file
setwd("~/Work/ETH/Projects/Nematodes/Figures/")
dev.copy(png, file = "stressplot.png")
dev.off()
```

Stress value of NMDS
```{r}
NemaNMDS$stress
```

## NMDS plot + vectors
Then, the NMDS was fitted with environmental vectors. 
```{r NMDS plot}
# Load environmental data, select rows based on ID
envfactors <- merge(Used_IDs, Data_raw, by = "ID" )

# select layers
envfactors<- envfactors[,c(
  'Annual_Precipitation',
  'Aridity_Index',
  'Sand_Content_15cm',
  'OrgCStockTHa_5to15cm',
  'NDVI',
  'Annual_Mean_Temperature',
  'Shannon_Index_1km',
  'Precipitation_Seasonality',
  'Temperature_Seasonality',
  'pHinHOX_15cm',
  "EVI",
  'Human_Development_Percentage')]

# envfit
fit <- envfit(NemaNMDS, envfactors, na.rm = T, permutations = 999)

## Plot NMDS 
## source https://oliviarata.wordpress.com/2014/04/17/ordinations-in-ggplot2/

# NMDS points
NMDS.data <- envfactors 
NMDS.data$NMDS1 <- NemaNMDS$points[ ,1] 
NMDS.data$NMDS2 <- NemaNMDS$points[ ,2] 
NMDS.data <- tibble::rowid_to_column(NMDS.data, "ID")
NMDS.data <- merge (NMDS.data, cluster_info, "ID")
NMDS.data <- melt(NMDS.data, c("NMDS1","NMDS2","Cluster"))

# Vectors
env.scores.nema <- as.data.frame(scores(fit, display = "vectors")) 
env.scores.nema <- cbind(env.scores.nema, env.variables = rownames(env.scores.nema)) 

col = colorRampPalette(c("#A9A9A9", "#C0C0C0", "#D3D3D3", "#696969")) 

NMDS <- ggplot(data = NMDS.data, aes(
  x = NMDS1,
  y = NMDS2)) + 
  geom_point(aes(color = factor(NMDS.data$Cluster), shape = factor(NMDS.data$Cluster)), size = 0.5) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line()) +
  scale_shape_manual(values = c(0,18,1,17), 
                     name = "Community type") +
    scale_color_manual(values = col(4),
                     name = "Community type") +
  geom_segment(data = env.scores.nema,
               aes(x = 0, 
                   xend = 4 * NMDS1, 
                   y = 0, 
                   yend = 4 * NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "black") +
  geom_text(data = env.scores.nema, 
            aes(x = 4.5 * NMDS1, 
                y = 4.5 * NMDS2, 
                label = env.variables),
            size = 3) +
  annotate("text", x = -0.9, y = 1.5, label = paste("Stress value = ", format(round(NemaNMDS$stress, 4), nsmall = 4)), size = 3)
NMDS

fit

#save to file
setwd("~/Work/ETH/Projects/Nematodes/Figures/")
ggsave("NMDS.pdf", plot = NMDS, device = "pdf", height =  10, width = 10, units = "cm", dpi = "retina", scale = 2)
```

