---
title: "Supplementary Methods - Biomass and Carbon budget calculationss"
author: "J. van den Hoogen & S. Giesen et al. 2019"
output:
  html_document:
    # code_folding: hide
    toc: yes
    toc_collapse: no
    toc_float: yes
    theme: spacelab
    fig_retina: 2
  github_document: default
---

```{r load packages, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
```
Initial values and assumptions
```{r Initial values and assumptions}
#Percentage of carbon (C) per fresh mass
cperw = 0.2 * 0.52

##1. Correct for the expected difference between CO2 evolved and O2 consumed using a respiratory quotient (RQ) of 0.95 
##Klekowski 1972 (doi: 10.1163/187529272x00665) reports a aO2 value of 1.40 nl O2 per hour. This can be converted to a aCO2 value of 2.43 using:
##aCO2 = mass.CO2 * (44 x 10^9)
a.O2 = 1.4
V.CO2 = 0.95 * a.O2

##2. Calculate the number of moles (m) of CO2
##We assume a temperature of 15 degrees Celcius.
t = 15
m = V.CO2 / (0.0821 * (273 + t) * 1E9)

##3. Convert from aO2 to aCO2 
a.CO2 = m * (44 * 1E9)

##4. Convert to daily rate, scale to relative mass of C in CO2. Divide by 1000 to convert from ng to ug.
a.CO2.daily = a.CO2 * 24 * (12/44) / 1000
```

Data source: 
For adult biomass and carbon budget calculations we use data from Nemaplex (http://nemaplex.ucdavis.edu/) and Mulder and Vonk 2011, doi:10.2307/23034835. For juveniles, we use data from Mulder and Vonk 2011. 

# Data import
```{r}
# Nemaplex data
setwd("~/Work/ETH/Projects/Nematodes/Calculations")
Nemaplex_data_raw <- read.csv("Family Ecophysiology qParameters.csv", stringsAsFactors = FALSE, fileEncoding = "latin1") 

# Mulder 2011 data
mulder_2011 <- read.csv("Mulder2011.csv")
mulder_data <- mulder_2011 %>% 
  select(trophic.group, Length, Width, Juv_Adult)

# Biome abundances
Biome_Abundances <- read.csv("20180827_Biome_Abundances.csv")

# Lowest estimates
numbers_min <- read.csv("20180912_Biome_Abundances_Minimum.csv") %>% 
  select(-Total_Number) %>% 
  gather(key, value, -Biome) %>% 
  rename(trophic.group = key) %>% 
  rename(Total.number = value)

# Highest estimates
numbers_max <- read.csv("20180912_Biome_Abundances_Maximum.csv") %>% 
  select(-Total_Number) %>% 
  gather(key, value, -Biome) %>% 
  rename(trophic.group = key) %>% 
  rename(Total.number = value)

adult_data_1 <- Nemaplex_data_raw %>% 
  select(cp.value, feeding.code, Length.micm, Width.micm) %>% 
  mutate(feeding.code = replace(feeding.code, feeding.code == "p", "Predators")) %>% 
  mutate(feeding.code = replace(feeding.code, feeding.code == "o", "Omnivores")) %>% 
  mutate(feeding.code = replace(feeding.code, feeding.code == "b", "Bacterivores")) %>% 
  mutate(feeding.code = replace(feeding.code, feeding.code == "h", "Herbivores")) %>% 
  mutate(feeding.code = replace(feeding.code, feeding.code == "f", "Fungivores")) %>% 
  filter(feeding.code != "e") %>% 
  filter(feeding.code != "d") %>% 
  filter(feeding.code != "") %>% 
  rename(trophic.group = feeding.code) %>% 
  rename(Length = Length.micm) %>% 
  rename(Width = Width.micm)

cp.values <- adult_data_1 %>% group_by(trophic.group) %>% summarise(cp.value = mean(cp.value)) 

mulder_data$trophic.group <- substr(mulder_data$trophic.group,1,4)  
adult_data_2 <- mulder_data %>%   
  mutate(trophic.group = replace(trophic.group, trophic.group == "bact", "Bacterivores")) %>% 
  mutate(trophic.group = replace(trophic.group, trophic.group == "fung", "Fungivores")) %>% 
  mutate(trophic.group = replace(trophic.group, trophic.group == "herb", "Herbivores")) %>% 
  mutate(trophic.group = replace(trophic.group, trophic.group == "omni", "Omnivores")) %>% 
  mutate(trophic.group = replace(trophic.group, trophic.group == "pred", "Predators")) %>% 
  filter(Juv_Adult == "adult")
adult_data_2 <- adult_data_2 %>% select(-Juv_Adult)
adult_data_2 <- merge(cp.values, adult_data_2[1:3], by = "trophic.group")

adult_data <- rbind(adult_data_1, adult_data_2)
```

# Bodymass calculations
## Adults
```{r}
#mass as function of length and width: W = (L ∙D^2)/(1.6 ∙ 10^6 )
adult_data$mass <- with(adult_data,(Length * Width * Width)/(1.6*1E6))

#Respiration rate in ug C per day as funciton of mass and aC02. b-value is set to 0.75
adult_data$Respiration.rate <- with(adult_data,(a.CO2.daily * ((adult_data$mass)^0.75)))

#Production rate in ug C per day as function of mass and cp value
adult_data$Production.rate <- with(adult_data, (cperw* ((adult_data$mass)/(12 * adult_data$cp.value))))

#Daily carbon budget in ug C; sum of respiration rate and production rate
adult_data$Carbon.budget <- with(adult_data, (adult_data$Respiration.rate + adult_data$Production.rate))

summary_adult <- adult_data %>% 
  group_by(trophic.group) %>% 
  summarise(n.measurements = n(), 
            mean.mass.ug = mean(mass),
            mean.daily.respiration.rate.ug = mean(Respiration.rate),
            mean.daily.production.rate.ug = mean(Production.rate),
            mean.daily.carbon.budget.ug = mean(Carbon.budget)) %>% 
  na.omit() 
summary_adult
```

## Juveniles
```{r}
juv_data <- mulder_2011 %>% 
  select(Taxonomy, trophic.group, Length, Width, Juv_Adult)

juv_data$trophic.group <- substr(juv_data$trophic.group,1,4)  
juv_data <- juv_data %>%   
  mutate(trophic.group = replace(trophic.group, trophic.group == "bact", "Bacterivores")) %>% 
  mutate(trophic.group = replace(trophic.group, trophic.group == "fung", "Fungivores")) %>% 
  mutate(trophic.group = replace(trophic.group, trophic.group == "herb", "Herbivores")) %>% 
  mutate(trophic.group = replace(trophic.group, trophic.group == "omni", "Omnivores")) %>% 
  mutate(trophic.group = replace(trophic.group, trophic.group == "pred", "Predators")) %>% 
  filter(Juv_Adult == "juveniles")

cp.values <- adult_data %>% group_by(trophic.group) %>% summarise(cp.value = mean(cp.value)) 

juv_data<- merge(cp.values, juv_data[1:5], by = "trophic.group")

#Mass as function of length and width: W = (L x D^2)/(1.6 x 10^6 )
juv_data$Mass <- with(juv_data, (Length * Width * Width)/(1.6*1E6))

#Respiration rate in ug C per day as function of mass. aCO2 is as calculated above b-value assumed at 0.75
juv_data$Respiration.rate <- with(juv_data,(a.CO2.daily * ((juv_data$Mass)^0.75)))

#Production rate in ug C per day as function of mass and cp value
juv_data$Production.rate <- with(juv_data, (cperw* ((juv_data$Mass)/(12 * juv_data$cp.value))))

#Daily carbon budget in ug C; sum of respiration rate and production rate
juv_data$Carbon.budget <- with(juv_data, (juv_data$Respiration.rate + juv_data$Production.rate))

summary_juv <- juv_data %>% 
  group_by(trophic.group) %>% 
  summarise(n.measurements = n(), 
            mean.mass.ug = mean(Mass),
            mean.daily.respiration.rate.ug = mean(Respiration.rate),
            mean.daily.production.rate.ug = mean(Production.rate),
            mean.daily.carbon.budget.ug = mean(Carbon.budget)) %>% 
  na.omit() 
summary_juv
```
# Number of nematodes per biome
The numbers per biome are outputted from the geospatial models.
```{r}
#Import calculated totals of individuals
numbers <- Biome_Abundances %>% 
  select(-Total_Number) %>% 
  gather(key, value, -Biome) %>% 
  rename(trophic.group = key) %>% 
  rename(Total.number = value)

#Total number of individuals, per trophic group
numbers_tot <- numbers %>% 
  group_by(trophic.group) %>% 
  summarise(tot.numb.indiv = sum(Total.number))

#Total number of individuals, per trophic group, per biome
numbers_biome <- numbers%>% group_by(trophic.group, Biome) %>% summarise(tot.numb.indiv = sum(Total.number)) 
```

## Per-trophic group sums
```{r}
#Total number of individuals, per trophic group
numbers_tot_print <- numbers_tot  %>% 
  rename("Trophic group" = 1) %>% 
  rename("Number of individuals" = 2) %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total")))
numbers_tot_print
```



## Relative abundances per biome
```{r}
#Relative abundance per biome
abundances <- Biome_Abundances %>%
  mutate(Biome = replace(Biome, Biome == 1, "Tropical Moist Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 2, "Tropical Dry Forest")) %>%
  mutate(Biome = replace(Biome, Biome == 3, "Tropical Coniferous Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 4, "Temperate Broadleaf Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 5, "Temperate Conifer Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 6, "Boreal Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 7, "Tropical Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 8, "Temperate Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 9, "Flooded Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 10, "Montane Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 11, "Tundra")) %>%
  mutate(Biome = replace(Biome, Biome == 12, "Mediterranean Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 13, "Deserts")) %>% 
  mutate(Biome = replace(Biome, Biome == 14, "Mangroves"))

abundances_biome <- abundances %>% select(-Total_Number)
abundances_biome[,2:6] <- prop.table(as.matrix(abundances_biome[,2:6]), 1)
abundances_biome[,2:6] <- round(select_if(abundances_biome[,2:6], is.numeric), 3)
abundances_biome

abundances_tot <- abundances %>% select(Biome, Total_Number) 
abundances_tot[,2] <- prop.table(as.matrix(abundances_tot[,2]), 2)
abundances_tot[,2] <- round(abundances_tot[,2], 3)
abundances_tot

rel.abundance.trophic <- format(round(prop.table(data.frame(as.list(apply(subset(abundances, select= c(-Biome, -Total_Number)), 2, sum)))), 3), nsmall = 3)
rel.abundance.trophic

abundances_final <- rbind(rel.abundance.trophic)

write.csv(abundances_final, "TableS5_rel_abundance_biome.csv")
```

# Carbon budget
## Adults
```{r}
#Adults
summary_tot_adult <- merge(summary_adult, numbers_tot, by = "trophic.group")
#Total biomass in Gt: mean mass per trophic group x individuals
summary_tot_adult$total.biomass.Gt <- with(summary_tot_adult, (summary_tot_adult$mean.mass.ug * summary_tot_adult$tot.numb.indiv)/1E21)
#Total biomass in Gt C: biomass x 20% fresh mass x carbon content 52% of dry mass
summary_tot_adult$total.Cbiomass.Gt <- with(summary_tot_adult, (summary_tot_adult$total.biomass.Gt * cperw))
#Total monthly respiration in Gt C: daily rate x individuals x 31 days
summary_tot_adult$total.mon.resp.Gt <- with(summary_tot_adult, (summary_tot_adult$mean.daily.respiration.rate.ug * summary_tot_adult$tot.numb.indiv * 31)/1E21)
#Total monthly production in Gt C: daily rate x individuals x 31 days
summary_tot_adult$total.mon.prod.Gt <- with(summary_tot_adult, (summary_tot_adult$mean.daily.production.rate.ug * summary_tot_adult$tot.numb.indiv * 31)/1E21)
#Total monthly carbon budget in Gt C: sum of respiration + production
summary_tot_adult$total.C.budget.Gt <- with(summary_tot_adult, (summary_tot_adult$total.mon.prod.Gt + summary_tot_adult$total.mon.resp.Gt))

#Final output
summary_tot_adult <- summary_tot_adult %>%  
  select(trophic.group, tot.numb.indiv, total.biomass.Gt, total.Cbiomass.Gt, total.mon.resp.Gt, total.mon.prod.Gt, total.C.budget.Gt) %>% 
  rename("Trophic group" = 1) %>% 
  rename("Computed individuals" = 2) %>% 
  rename("Fresh biomass (Mt)" = 3) %>% 
  rename("Biomass (Mt C)" = 4) %>% 
  rename("Monthly respiration (Mt C)" = 5) %>% 
  rename("Monthly production (Mt C)" = 6) %>% 
  rename("Monthly carbon budget (Mt C)" = 7) %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total")))

summary_tot_adult[,3:7] <- summary_tot_adult[,3:7] * 1000
summary_tot_adult[,-1] <- round(select_if(summary_tot_adult, is.numeric), 2)
summary_tot_adult[,2] <- formatC(summary_tot_adult[,2], format = "e", digits = 2)
summary_tot_adult
```

## Juveniles
```{r}
#Juveniles
summary_tot_juv <- merge(summary_juv, numbers_tot, by = "trophic.group")
#Total biomass in Gt: mean mass per trophic group x individuals
summary_tot_juv$total.biomass.Gt <- with(summary_tot_juv, (summary_tot_juv$mean.mass.ug * summary_tot_juv$tot.numb.indiv)/1E21)
#Total biomass in Gt C: biomass x 20% fresh mass x carbon content 52% of dry mass
summary_tot_juv$total.Cbiomass.Gt <- with(summary_tot_juv, (summary_tot_juv$total.biomass.Gt * cperw))
#Total monthly respiration in Gt C: daily rate x individuals x 31 days
summary_tot_juv$total.mon.resp.Gt <- with(summary_tot_juv, (summary_tot_juv$mean.daily.respiration.rate.ug * summary_tot_juv$tot.numb.indiv * 31)/1E21)
#Total monthly production in Gt C: daily rate x individuals x 31 days
summary_tot_juv$total.mon.prod.Gt <- with(summary_tot_juv, (summary_tot_juv$mean.daily.production.rate.ug * summary_tot_juv$tot.numb.indiv * 31)/1E21)
# #Total monthly carbon budget in Gt C: sum of respiration + production
summary_tot_juv$total.C.budget.Gt <- with(summary_tot_juv, (summary_tot_juv$total.mon.prod.Gt + summary_tot_juv$total.mon.resp.Gt))

summary_tot_juv <- summary_tot_juv %>%  
  select(trophic.group, tot.numb.indiv, total.biomass.Gt, total.Cbiomass.Gt, total.mon.resp.Gt, total.mon.prod.Gt, total.C.budget.Gt) %>% 
  rename("Trophic group" = 1) %>% 
  rename("Computed individuals" = 2) %>% 
  rename("Fresh biomass (Mt)" = 3) %>% 
  rename("Biomass (Mt C)" = 4) %>% 
  rename("Monthly respiration (Mt C)" = 5) %>% 
  rename("Monthly production (Mt C)" = 6) %>% 
  rename("Carbon budget (Mt C)" = 7) %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total")))

summary_tot_juv[,3:7] <- summary_tot_juv[,3:7] * 1000
summary_tot_juv[,-1] <- round(select_if(summary_tot_juv, is.numeric), 2)
summary_tot_juv[,2] <- formatC(summary_tot_juv[,2], format = "e", digits = 2)
```

## Total carbon budget
```{r}
# Final table of biomass and carbon budget

#Assuming a relative proportion of 30% adults and 70% juveniles
adult_frac <- 0.30
juv_frac <- 0.70

#Multiply adult and juvenile summaries by relative proportions
summary_tot_final <- summary_tot_juv[,1:2]
summary_tot_final[3:7] <- ((adult_frac * summary_tot_adult[3:7]) + (juv_frac * summary_tot_juv[3:7])) 
summary_tot_final[3:7] <- round(select_if(summary_tot_final[3:7], is.numeric), 2)
summary_tot_final

#Write to file
setwd("~/Work/ETH/Projects/Nematodes/Calculations")
write.csv(summary_tot_final, "Table1_carbon_budget_2.csv")
```

## outer bounds (i.e. ± values) in our estimates 
```{r}
#to calculate the spread of the data, we take the minimum and maximum predicted values across all ensembled models
numbers_tot_min <- numbers_min %>% 
  group_by(trophic.group) %>% 
  summarise(tot.numb.indiv = sum(Total.number))

numbers_tot_max <- numbers_max %>% 
  group_by(trophic.group) %>% 
  summarise(tot.numb.indiv = sum(Total.number))

numbers_tot_min_print <- numbers_tot_min %>% 
  rename("Trophic group" = 1) %>% 
  rename("Number of individuals" = 2) %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total")))

numbers_tot_max_print <- numbers_tot_max %>% 
  rename("Trophic group" = 1) %>% 
  rename("Number of individuals" = 2) %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total")))

variation <- numbers_tot_print[,1]
#variation$avg <- (((numbers_tot_max_print$"Number of individuals" - numbers_tot_print$"Number of individuals") + (numbers_tot_print$"Number of individuals" - numbers_tot_min_print$"Number of individuals"))/2)
variation$mean.min <- (numbers_tot_print$"Number of individuals" - numbers_tot_min_print$"Number of individuals")
#variation$max.mean <- (numbers_tot_max_print$"Number of individuals" - numbers_tot_print$"Number of individuals") 

#Adults
summary_tot_min_adult <- merge(summary_adult, numbers_tot_min, by = "trophic.group")
#Total biomass in Gt: mean mass per trophic group x individuals
summary_tot_min_adult$total.biomass.Gt <- with(summary_tot_min_adult, (summary_tot_min_adult$mean.mass.ug * summary_tot_min_adult$tot.numb.indiv)/1E21)
#Total biomass in Gt C: biomass x 20% fresh mass x carbon content 52% of dry mass
summary_tot_min_adult$total.Cbiomass.Gt <- with(summary_tot_min_adult, (summary_tot_min_adult$total.biomass.Gt * cperw))
#Total monthly respiration in Gt C: daily rate x individuals x 31 days
summary_tot_min_adult$total.mon.resp.Gt <- with(summary_tot_min_adult, (summary_tot_min_adult$mean.daily.respiration.rate.ug * summary_tot_min_adult$tot.numb.indiv * 31)/1E21)
#Total monthly production in Gt C: daily rate x individuals x 31 days
summary_tot_min_adult$total.mon.prod.Gt <- with(summary_tot_min_adult, (summary_tot_min_adult$mean.daily.production.rate.ug * summary_tot_min_adult$tot.numb.indiv * 31)/1E21)
#Total monthly carbon budget in Gt C: sum of respiration + production
summary_tot_min_adult$total.C.budget.Gt <- with(summary_tot_min_adult, (summary_tot_min_adult$total.mon.prod.Gt + summary_tot_min_adult$total.mon.resp.Gt))

#Final output
summary_tot_min_adult <- summary_tot_min_adult %>%  
  select(trophic.group, tot.numb.indiv, total.biomass.Gt, total.Cbiomass.Gt, total.mon.resp.Gt, total.mon.prod.Gt, total.C.budget.Gt) %>% 
  rename("Trophic group" = 1) %>% 
  rename("Computed individuals" = 2) %>% 
  rename("Fresh biomass (Mt)" = 3) %>% 
  rename("Biomass (Mt C)" = 4) %>% 
  rename("Monthly respiration (Mt C)" = 5) %>% 
  rename("Monthly production (Mt C)" = 6) %>% 
  rename("Monthly carbon budget (Mt C)" = 7) %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total")))

summary_tot_min_adult[3:7] <- summary_tot_min_adult[3:7] * 1000
summary_tot_min_adult[,-1] <- round(select_if(summary_tot_min_adult, is.numeric), 3)
summary_tot_min_adult[,2] <- formatC(summary_tot_min_adult[,2], format = "e", digits = 3)

#Juveniles
summary_tot_min_juv <- merge(summary_juv, numbers_tot_min, by = "trophic.group")
#Total biomass in Gt: mean mass per trophic group x individuals
summary_tot_min_juv$total.biomass.Gt <- with(summary_tot_min_juv, (summary_tot_min_juv$mean.mass.ug * summary_tot_min_juv$tot.numb.indiv)/1E21)
#Total biomass in Gt C: biomass x 20% fresh mass x carbon content 52% of dry mass
summary_tot_min_juv$total.Cbiomass.Gt <- with(summary_tot_min_juv, (summary_tot_min_juv$total.biomass.Gt * cperw))
#Total monthly respiration in Gt C: daily rate x individuals x 31 days
summary_tot_min_juv$total.mon.resp.Gt <- with(summary_tot_min_juv, (summary_tot_min_juv$mean.daily.respiration.rate.ug * summary_tot_min_juv$tot.numb.indiv * 31)/1E21)
#Total monthly production in Gt C: daily rate x individuals x 31 days
summary_tot_min_juv$total.mon.prod.Gt <- with(summary_tot_min_juv, (summary_tot_min_juv$mean.daily.production.rate.ug * summary_tot_min_juv$tot.numb.indiv * 31)/1E21)
# #Total monthly carbon budget in Gt C: sum of respiration + production
summary_tot_min_juv$total.C.budget.Gt <- with(summary_tot_min_juv, (summary_tot_min_juv$total.mon.prod.Gt + summary_tot_min_juv$total.mon.resp.Gt))

summary_tot_min_juv <- summary_tot_min_juv %>%  
  select(trophic.group, tot.numb.indiv, total.biomass.Gt, total.Cbiomass.Gt, total.mon.resp.Gt, total.mon.prod.Gt, total.C.budget.Gt) %>% 
  rename("Trophic group" = 1) %>% 
  rename("Computed individuals" = 2) %>% 
  rename("Fresh biomass (Mt)" = 3) %>% 
  rename("Biomass (Mt C)" = 4) %>% 
  rename("Monthly respiration (Mt C)" = 5) %>% 
  rename("Monthly production (Mt C)" = 6) %>% 
  rename("Carbon budget (Mt C)" = 7) %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total")))

summary_tot_min_juv[3:7] <- summary_tot_min_juv[3:7] * 1000
summary_tot_min_juv[,-1] <- round(select_if(summary_tot_min_juv, is.numeric), 3)
summary_tot_min_juv[,2] <- formatC(summary_tot_min_juv[,2], format = "e", digits = 3)

#Multiply adult and juvenile summaries by relative proportions
summary_tot_min_final <- summary_tot_min_juv[,1:2]
summary_tot_min_final[3:7] <- ((adult_frac * summary_tot_min_adult[3:7]) + (juv_frac * summary_tot_min_juv[3:7]))

variation <- variation %>% 
  rename("Trophic group" = 1) %>% 
  rename("Number ±" = 2)
variation$'Fresh biomass ±' <- summary_tot_final$`Fresh biomass (Mt)` - summary_tot_min_final$`Fresh biomass (Mt)`
variation$"Biomass C ±" <- summary_tot_final$`Biomass (Mt C)` - summary_tot_min_final$`Biomass (Mt C)`
variation$"Respiration ±" <- summary_tot_final$`Monthly respiration (Mt C)` - summary_tot_min_final$`Monthly respiration (Mt C)`
variation$"Production ±" <- summary_tot_final$`Monthly production (Mt C)` - summary_tot_min_final$`Monthly production (Mt C)`
variation$"Budget ±" <- summary_tot_final$`Monthly carbon budget (Mt C)` - summary_tot_min_final$`Monthly carbon budget (Mt C)`
variation

write.csv(variation, "variation.csv")

```

#Biomass per biome
```{r}
#Total biomass per biome, per trophic group

#adult
summary_biome_adult <- merge(summary_adult, numbers_biome, by = "trophic.group")
summary_biome_adult$total.mon.resp.Gt <- with(summary_biome_adult, (summary_biome_adult$mean.daily.respiration.rate.ug * summary_biome_adult$tot.numb.indiv * 31)/1E21)
summary_biome_adult$total.biomass.Gt <- with(summary_biome_adult, (summary_biome_adult$mean.mass.ug * summary_biome_adult$tot.numb.indiv)/1E21)
summary_biome_adult$total.biomass.C.Gt <- with(summary_biome_adult, (summary_biome_adult$total.biomass.Gt * cperw * 1000))
summary_biome_adult$total.mon.prod.Gt <- with(summary_biome_adult, (summary_biome_adult$mean.daily.production.rate.ug * summary_biome_adult$tot.numb.indiv * 31)/1E21)
summary_biome_adult$total.C.budget.Gt <- with(summary_biome_adult, (summary_biome_adult$total.mon.prod.Gt + summary_biome_adult$total.mon.resp.Gt))


biomesum_adult <- summary_biome_adult %>% group_by(Biome, trophic.group) %>% summarise(sum(total.biomass.C.Gt)) %>%
  spread(key = trophic.group, value = 'sum(total.biomass.C.Gt)') 
biomesum_adult <- transform(biomesum_adult, Total.biomass=rowSums(subset(biomesum_adult, select= -Biome))) %>% 
  mutate(Biome = replace(Biome, Biome == 1, "Tropical Moist Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 2, "Tropical Dry Forest")) %>%
  mutate(Biome = replace(Biome, Biome == 3, "Tropical Coniferous Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 4, "Temperate Broadleaf Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 5, "Temperate Conifer Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 6, "Boreal Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 7, "Tropical Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 8, "Temperate Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 9, "Flooded Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 10, "Montane Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 11, "Tundra")) %>%
  mutate(Biome = replace(Biome, Biome == 12, "Mediterranean Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 13, "Deserts")) %>% 
  mutate(Biome = replace(Biome, Biome == 14, "Mangroves")) %>% 
  mutate(Biome = replace(Biome, Biome == 105, "Total"))

# juveniles
summary_biome_juv <- merge(summary_juv, numbers_biome, by = "trophic.group")
summary_biome_juv$total.mon.resp.Gt <- with(summary_biome_juv, (summary_biome_juv$mean.daily.respiration.rate.ug * summary_biome_juv$tot.numb.indiv * 31)/1E21)
summary_biome_juv$total.biomass.Gt <- with(summary_biome_juv, (summary_biome_juv$mean.mass.ug * summary_biome_juv$tot.numb.indiv)/1E21)
summary_biome_juv$total.biomass.C.Gt <- with(summary_biome_juv, (summary_biome_juv$total.biomass.Gt * cperw * 1000))
summary_biome_juv$total.mon.prod.Gt <- with(summary_biome_juv, (summary_biome_juv$mean.daily.production.rate.ug * summary_biome_juv$tot.numb.indiv * 31)/1E21)
summary_biome_juv$total.C.budget.Gt <- with(summary_biome_juv, (summary_biome_juv$total.mon.prod.Gt + summary_biome_juv$total.mon.resp.Gt))

biomesum_juv <- summary_biome_juv %>% group_by(Biome, trophic.group) %>% summarise(sum(total.biomass.C.Gt)) %>%
  spread(key = trophic.group, value = 'sum(total.biomass.C.Gt)') 
biomesum_juv <- transform(biomesum_juv, Total.biomass=rowSums(subset(biomesum_juv, select= -Biome))) %>% 
  mutate(Biome = replace(Biome, Biome == 1, "Tropical Moist Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 2, "Tropical Dry Forest")) %>%
  mutate(Biome = replace(Biome, Biome == 3, "Tropical Coniferous Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 4, "Temperate Broadleaf Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 5, "Temperate Conifer Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 6, "Boreal Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 7, "Tropical Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 8, "Temperate Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 9, "Flooded Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 10, "Montane Grasslands")) %>%
  mutate(Biome = replace(Biome, Biome == 11, "Tundra")) %>%
  mutate(Biome = replace(Biome, Biome == 12, "Mediterranean Forests")) %>%
  mutate(Biome = replace(Biome, Biome == 13, "Deserts")) %>% 
  mutate(Biome = replace(Biome, Biome == 14, "Mangroves")) %>% 
  mutate(Biome = replace(Biome, Biome == 105, "Total"))

biomesum <- biomesum_adult[,1, drop = FALSE]
biomesum[2:7] <- ((adult_frac * biomesum_adult[2:7]) + (juv_frac * biomesum_juv[2:7]))

biomesum_print <- biomesum %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total"))) %>% 
  rename("Bacterivores (Mt C)" = 2) %>% 
  rename("Fungivores (Mt C)" = 3) %>% 
  rename("Herbivores (Mt C)" = 4) %>% 
  rename("Omnivores (Mt C)" = 5) %>% 
  rename("Predators (Mt C)" = 6) %>% 
  rename("Total biomass (Mt C)" = 7) 
biomesum_print[-1] <- format(round(biomesum_print[-1], 3), nsmall = 5)
biomesum_print

write.csv(biomesum_print, "TableS6_biomass.csv")
```

#Relative biomass of each trophic group
```{r}
#Relative biomass of each trophic group 
rel.biomass.trophic <- format(round(prop.table(data.frame(as.list(apply(subset(biomesum, select= c(-Biome, -Total.biomass)), 2, sum)))), 3), nsmall = 3)
rel.biomass.trophic
```

#Relative biomass of each trophic group per biome
```{r}
#Relative biomass of each trophic group per biome
rel.biomass.biome <- as.data.frame(prop.table(as.matrix(biomesum[2:6]),1))
rel.biomass.biome <- cbind(biomesum$Biome, format(round(rel.biomass.biome, 3), nsmall = 3)) %>% 
  rename(Biome = 1)

rel.biomass.trophic$Biome <- "Total"
rel.biomass <- rbind(rel.biomass.biome,rel.biomass.trophic)
```

```{r}
#relative biomass per biome
rel.biomass.sum <- as.data.frame(prop.table(as.matrix(subset(biomesum, select = Total.biomass))),1) %>% 
  bind_rows(summarise_all(., funs(if(is.numeric(.)) sum(.) else "Total"))) %>% 
  rename("Across trophic groups - fraction of total biomass" = 1)

rel.biomass.final <- cbind(rel.biomass, format(round(rel.biomass.sum, 3), nsmall = 3)) %>% 
  rename("Bacterivores - fraction of total" = 2) %>% 
  rename("Fungivores - fraction of total" = 3) %>% 
  rename("Herbivores - fraction of total" = 4) %>% 
  rename("Omnivores - fraction of total" = 5) %>% 
  rename("Predators - fraction of total" = 6)
rel.biomass.final

write.csv(rel.biomass.final, "TableS7_rel_biomass.csv")
```

#Comparing nematodes to humans
```{r}
#total nematode biomass (freshweight) in Gt:
tot_nema_biomass <- summary_tot_final$`Fresh biomass (Gt)`[6]

#Human biomass is calculated using assumptions from Bar-On et al. 2018 PNAS 
#human population from 2015 UN data
human_pop <- 7383008.82 * 1000 #https://esa.un.org/unpd/wpp/Download/Standard/Population/

#human population versus nematode population in billions
cat("For each human on Earth,", format(round((numbers_tot_print$`Number of individuals`[6]/human_pop)/1E9, 1), nsmall = 1), "billion nematodes are present.")

#Average human weight assumed at 50 kg by Hern et al. 1999 https://link.springer.com/article/10.1023/A:1022153110536
human_weight <- 50
human_biomass <- (human_pop * human_weight)/1E12

#nematode biomass as fraction of human biomass
cat("Total soil nematode biomass represents", format(round((tot_nema_biomass/human_biomass*100), 1)),"% of total human biomass")
```

